\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb}
\usepackage{enumitem}
\usepackage{xspace}
\usepackage{tcolorbox}
\usepackage[colorlinks=true,linkcolor=blue,citecolor=blue,urlcolor=blue]{hyperref}

\newcommand{\tool}{\textsc{SkillRACE}\xspace}
\newcommand{\bench}{\textsc{SkillStress}\xspace}

\title{\tool: Implementation Pipeline\\
\large Reasoning-Augmented Concolic Execution for Coding-Agent Skills\\
\normalsize Engineering design document --- what to build, in what order, and why it produces results}
\author{}
\date{June 2026}

\begin{document}
\maketitle

\section{What this system is, in one paragraph}

A coding-agent \emph{skill} is reusable procedural guidance (a \texttt{SKILL.md}
plus optional scripts) that an agent consults to do a task---fix a failing
test, rebase a branch, scaffold a component. Skills are trusted after a handful
of runs and tested almost not at all. \tool tests them the way concolic
execution tests code: it runs the skill, lifts each run into a sequence of
\emph{episodes} (coherent units of work, each with a purpose and an
observation-grounded outcome), merges episodes into a growing \emph{behavior
tree}, reads the agent's reasoning at each branch as the \emph{condition} that
sent it one way rather than another, and then \emph{negates or mutates} those
conditions to synthesize new (prompt, environment) inputs that drive the skill
down branches it has not yet taken. Each new run folds back into the tree, which
reveals the next unexplored branch, and so on. The tree is both the byproduct
and the driver of the search---exactly as the path tree is in classical
concolic execution. The payoff is twofold: (i)~a map of how the skill's guidance
actually maps onto execution, showing which parts of the skill remain
unexplored, and (ii)~a stream of targeted tests that expose where the skill
violates its intended behavior. Correctness of each run is judged by
\emph{properties}---some fixed formulas, some natural-language specifications
compiled per task into executable checks (specification by example)---not by
asking a model ``did this go well?''

\paragraph{The core conceptual move (the part that makes it publishable).}
Episodes are the \emph{basic blocks}; the agent's reasoning between them is the
\emph{branch condition}; hypothesizing over those conditions (negating,
mutating) tells us how the skill maps onto the execution trajectory, which
regions remain unexplored, and how to synthesize new interesting tests. This is
the concolic-execution idea lifted from source code to agent behavior: there the
branch condition is a path constraint read from the code; here it is a
natural-language condition read from the agent's reasoning.

\paragraph{What this document is.} The engineering plan: components, data
formats, the per-stage algorithm, the baselines, and the build order. The formal
treatment (soundness of the abstraction, etc.) lives in a separate document; here
the emphasis is on what to build and why it is realistic and produces results.

\paragraph{Models.} A single model is used for \emph{every} model-driven step
(segmentation, summarization, merge decisions, guard extraction, specification
compilation). We do not mix models across steps; instead we \emph{ablate} over
the choice of model (swap it for a stronger/weaker one and report how the
pipeline's numbers move). This keeps the design and the calibration uniform: we
calibrate one model per judgment type, and the model-ablation shows the whole
pipeline's sensitivity to it. The agent under test is whatever agent the skill
targets and is independent of this choice.

\section{System overview and cost structure}

The pipeline has six components. Five are cheap (plain code or a small fast
model); one---running the skill's agent under test---is expensive (minutes and
real money per run). \textbf{Every design decision follows from this asymmetry:
spend agent runs only on inputs we have already validated, and do everything else
with cheap models or code.}

\begin{enumerate}[leftmargin=1.5em, itemsep=2pt]
  \item \textbf{Runner} --- executes a skill in Docker on an input $(x, E_0)$,
  logs the full trace. \emph{(expensive: the agent under test)}
  \item \textbf{Episode segmenter} --- splits a trace into episodes and labels
  each with a purpose. \emph{(cheap model)}
  \item \textbf{Outcome \& summary} --- assigns each episode an
  observation-grounded outcome and a short canonical summary. \emph{(cheap model
  + code)}
  \item \textbf{Tree builder} --- merges episodes into the behavior tree;
  decides where runs share structure and where they branch. \emph{(cheap model
  for merge decisions; code for the tree)}
  \item \textbf{Guard extractor \& test synthesizer} --- reads the reasoning at
  a branch, distills the condition, mutates it, drafts a new $(x, E_0)$, and
  validates it before any agent run. \emph{(cheap model + code validator)}
  \item \textbf{Property checker} --- evaluates correctness properties over a
  run. \emph{(code; cheap model only to compile NL specs into checks)}
\end{enumerate}

Loop: seed the tree with a few runs $\to$ pick an unexplored branch $\to$
mutate its condition $\to$ synthesize and validate a new input $\to$ run the
agent $\to$ fold the trace back in $\to$ check properties $\to$ repeat until the
run budget is spent.

\section{Component 1: The runner and trace format}

\paragraph{What it does.} Given $(x, E_0)$---a prompt and a Docker
initialization script that builds the starting environment---it runs the skill's
agent at temperature~0 inside the container and records everything.

\paragraph{Trace format (frozen early; everything downstream consumes only
this).} The runner records the agent framework's raw session log; a
\emph{deterministic renderer} (no model) projects it into the globally-numbered
view every downstream component consumes --- one entry per tool call:
\begin{tcolorbox}[colback=gray!5,boxrule=0.4pt]
\small
\texttt{\{ call: int, reasoning: str,\ \ \ \ \ \ // assistant thinking before the call}\\
\texttt{\ \ tool: str, args: \{...\},\ \ \ \ \ \ \ // the tool call}\\
\texttt{\ \ observation: str \}\ \ \ \ \ \ \ \ \ \ \ \ // raw tool output}
\end{tcolorbox}

\noindent Final repository / container state is read directly at
property-checking time (Section~\ref{sec:properties}); we do not store a
per-step environment snapshot.

\paragraph{Why this is realistic.} Coding skills are short (procedural guidance
for one task), so traces are tens of steps, not thousands. The runner is a thin
wrapper around the agent framework plus Docker; the only real work is the
per-skill environment setup (sourcing skills with clean, Dockerizable tasks is
the real labor, not the runner). Use a hard
\textbf{step/turn cap} per run---needed for reproducibility and to stop a runaway
agent from burning budget.

\section{Component 2: Episode segmentation}
\label{sec:segment}

\paragraph{What an episode is.} A contiguous run of steps that together pursue
one purpose: ``locate the failing test,'' ``edit the implementation,'' ``run the
tests,'' ``verify the endpoint.'' An episode is the \textbf{basic block} of the
behavior tree---internally it may be one tool call or eight, but it is treated as
a unit with a purpose (intent) and a result (outcome).

\paragraph{How we split (and the principle behind it).} \textbf{Boundaries are
read from the agent's reasoning, because the agent announces its goal switches}
(``Now let me verify the endpoint works,'' ``First I'll check whether the tests
pass''). The segmenter is a cheap-model call over the trace that emits episode
boundaries, each with an \emph{intent} (a short purpose phrase) and an
\emph{evidence pointer} (the index of the reasoning turn that announced the
episode). Output is strict JSON, mechanically validated: episodes partition the
steps in order; every evidence index is in range. One retry, then the trace is
flagged unsegmentable and counted (expected rare).

\paragraph{Mechanics.} The renderer's numbered trace plus a baked-in worked
example goes to the segmenter with a \emph{soft episode-count target}, a smooth
saturating function of the tool-call count ($D = 3 + N/50$, target $= N/D$) that
discourages under-splitting without dictating boundaries. Spans are validated
mechanically (they must partition the calls in order; boundaries may only fall
where a reasoning turn exists); one repair attempt with the violation fed back,
then the trace is flagged unsegmentable and counted (expected rare). Each
episode carries \emph{intent}, \emph{what it did}, and \emph{outcome} --- the
summarization of Section~\ref{sec:summary} is emitted in the same pass. Traces
that exceed one context window are handled by an agentic variant of the same
segmenter that pages through the rendered trace file itself under identical
rules.

\paragraph{Faithfulness/quality knob.} Segmentation is a model judgment, so it is
a calibrated component: hand-segment $\sim$100 traces and measure boundary
agreement (F1) and intent-label agreement separately---a boundary error shifts
two adjacent episodes, a label error corrupts one. These numbers go in the
evaluation.

\section{Component 3: Episode summarization}
\label{sec:summary}

\paragraph{What it does.} One model call per episode turns the episode's raw
steps---its tool calls, its reasoning, and its tool outputs---into a short
\textbf{structured summary} following a fixed template: \emph{what was attempted,
on what, and with what result}. Example: \emph{``ran the project test suite; it
failed with an ImportError in the auth module.''} The summary is open in content
(the model captures whatever actually happened) but bounded in form (the
template), so two episodes can later be compared for ``same situation'' without
either pre-bucketing them into a rigid enum or leaving them as unmatchable
free text.

\paragraph{The result \emph{is} the outcome, and it is read from this episode's
tool outputs.} There is no separate ``outcome'' stage and no fixed outcome
grammar---the ``with what result'' part of the summary is the outcome. The one
rule that matters: that result is read from the episode's \emph{tool outputs}
(the exit code, the printed text, the HTTP status), \textbf{not} from the agent's
own narration of how it went. Agents declare false victory; if the recorded
result came from the agent saying ``tests pass,'' the skill would be grading
itself, and a run that ignored a real failure would be recorded as a success. By
sourcing the result from what the test command actually printed, a divergence
between \emph{what happened} (observation) and \emph{what the agent thought
happened} (its later reasoning) is preserved---and that divergence is itself a
bug signal (see Section~\ref{sec:guards}). In practice the model can read exit
codes and output directly; the template just asks it to ground the ``result''
field in the observation, not the reasoning.

\paragraph{Quality knob.} Summarization is a model judgment, so it is calibrated:
on a sample of episodes, check that the summary's result field agrees with the
ground-truth result of the tool outputs.

\section{Component 4: The behavior tree (online merging)}
\label{sec:tree}

\paragraph{What the tree is.} A growing structure whose nodes are
\emph{situations the skill reaches} and whose edges are \emph{what it did next}.
It is a \textbf{concolic search tree}: built incrementally as runs come in, and
used at each step to decide what to explore next. There is no ``collect all runs
then build''---later runs do not exist until the current tree has proposed them.

\paragraph{How a run folds in (the merge).} Walk the new run's episode sequence
from the root:
\begin{enumerate}[leftmargin=1.5em, itemsep=1pt]
  \item At the current node, compare the run's next episode against the current
  node's \emph{children}. Matching is on \textbf{purpose} (the episode's intent;
  how it was done is secondary context): a cheap-model similarity judgment, not
  string equality. \textbf{The outcome does not enter the match}---it is recorded
  on the \emph{edge} taken into the node (together with the opening reasoning of
  the next episode), which is exactly where the guard will read it
  (Section~\ref{sec:guards}).
  \item \textbf{Match} $\Rightarrow$ merge into that child: the node's purpose is
  \textbf{rewritten to generalize} over all its members (broadening only, so a
  node never disowns a run it already represents), and the episode's concrete
  tactic is recorded as one of the node's \emph{ways} (similar tactics collapse,
  genuinely different ones accumulate as variants).
  \item \textbf{No match} $\Rightarrow$ create a new child (the run diverges
  here). A node with $\ge 2$ children is a \textbf{branch}---the place a guard
  lives.
\end{enumerate}
Two runs that reach the same situation with \emph{different outcomes} therefore
share the node but carry different outcome labels on their out-edges---the
divergence, if it matters, appears one step later as a branch, with the differing
outcomes sitting on its in-edges as the guard's first signal.

\paragraph{Why similarity-merging, and why it is safe here.} A fixed outcome enum
would crush the diversity of real episodes; pure string matching would never
merge two phrasings of the same situation. Similarity-merging lets the model
judge ``same purpose'' with full nuance. The risk---that a similarity judgment
is not perfectly consistent---is contained two ways: (i)~merge decisions are made
at temperature~0 and \textbf{cached}, so the same episode pair always gets the
same verdict within a campaign; and (ii)~a \emph{wrong} merge is self-limiting:
if two genuinely different situations were merged, their next episodes diverge,
the point presents as a branch, and the guard extracted there names the
distinguishing condition---the search routes around the error rather than being
corrupted by it. Merge errors degrade search guidance temporarily; they do not
silently corrupt results.

\paragraph{Determinism, honestly.} The build is a deterministic \emph{procedure}
(same seeds + same selection policy + temperature~0 + caching $\Rightarrow$ same
campaign), not an order-invariant abstraction. We report build stability
empirically: re-run a campaign under a different random seed for the selection
policy and measure tree agreement.

\paragraph{What the tree gives us.} (i)~\textbf{Coverage}: how many distinct
situations/branches the skill's behavior occupies---and, by comparison against
the skill's documented steps, which parts of the skill remain unexercised.
(ii)~\textbf{Branch frontier}: the set of branches whose alternatives we have not
yet tried---the worklist for test synthesis.

\section{Component 5: Guards and test synthesis}
\label{sec:guards}

\paragraph{Where a guard lives, and the two distinct signals it uses.} A guard
sits at a \textbf{branch}---a point where runs that reached the same situation
went different ways---and is the \emph{condition} that explains the split. It is
read from two things, and it is worth being precise that these are
\emph{different signals with different sources}, because they are easy to
conflate:
\begin{itemize}[leftmargin=1.5em, itemsep=1pt]
  \item the \textbf{outcome of the episode that just finished}---this is the
  ``result'' field of that episode's summary (Section~\ref{sec:summary}), read
  from its \emph{tool outputs}; and
  \item the \textbf{opening reasoning of the next episode} (``the endpoint
  returned 500, so I'll check the logs'')---read from the agent's \emph{reasoning
  text}.
\end{itemize}
The outcome tells us \emph{what state the agent was in}; the next reasoning tells
us \emph{why it chose the branch it did, given that state}. They are not the same
source, even though the agent's reasoning usually \emph{restates} the
outcome---and that is the subtle point: we take the outcome from the observation
(Component~3), not from the reasoning. When the two disagree---the test
\emph{observably} failed but the agent's next reasoning proceeds as if it
passed---that disagreement is not noise to reconcile; it is a bug signal (the
agent ignored a real failure), and we record it. The guard is \emph{not} read
from the internal tool calls of an episode---those are tactics the tree abstracts
over, which is exactly why two runs that reached the same outcome by different
tactics present the same situation to the branch.

\paragraph{This is the conceptual core.} A classical program hands the tester its
branch conditions in source code. An agent has no source code---its reasoning
trace is the nearest thing: it is where the agent states why it branched. We
treat each stated reason as a \emph{candidate} branch condition, distilled into
a checkable form, and we \emph{test} it by mutation: if changing the condition
does not change the behavior, the stated reason was not the real cause (the loop
detects this and adjusts). Reasoning proposes the branch condition; execution
confirms or refutes it.

\paragraph{Guard extraction.} When a branch exists, a model call takes the
outcomes + opening reasonings of the diverging sides, plus a \textbf{diff} of
their initial environments/observations, and produces the guard as a short
condition \emph{plus, wherever possible, an executable grounding}: a concrete
check such as \texttt{test -f <path>}, ``exit code $\ne$ 0'', or ``output
contains \texttt{ModuleNotFoundError}''. Guards that ground to an executable
check are \emph{cheap to satisfy and verify}; guards that stay natural-language
are evaluated by a (calibrated) model when needed.

\paragraph{Test synthesis = negate or mutate the guard.} To explore an
unexplored branch, take a guard and drive it to a \emph{different} value than
observed:
\begin{itemize}[leftmargin=1.5em, itemsep=1pt]
  \item \textbf{Negate} (binary guard): the test failed with an assertion error
  $\to$ make it fail with something else / not fail.
  \item \textbf{Mutate} (multi-valued guard): the failure was an assertion error
  $\to$ target an import error, a timeout, a collection error---each a sibling
  region.
\end{itemize}
A cheap-model \textbf{generator} drafts a candidate $(x^*, E_0^*)$: a prompt plus
a Docker init script intended to satisfy the path into the branch \emph{and} the
mutated guard.

\paragraph{Property-guided targeting.} Which guard to mutate, and toward which
value, is not chosen blindly: the selection step is shown the skill's
\emph{correctness properties} (Section~\ref{sec:properties}) alongside the
frontier and asked for the feasible (guard, mutation) pair whose exploration is
most likely to drive the skill toward a property violation---e.g.\ prefer ``the
failing test is an import error'' $\to$ ``the failing test is an
\emph{environment} error'' when a test-integrity property could plausibly break
there. This turns exploration into \emph{directed falsification}: the frontier
supplies what is reachable, the properties supply what is worth reaching. The
choice is a targeting heuristic only---it biases the search; correctness verdicts
still come exclusively from the property checker.

\paragraph{Validation BEFORE spending an agent run (the key efficiency move).}
We only target guards that are decidable from the \textbf{initial setup}
($E_0$-guards: ``the repo lacks file X,'' ``the failing test is an import
error,'' ``dependencies are missing''), plus guards causally forced by $E_0$.
For such a guard, the \textbf{validator builds the candidate container and checks
the condition by running plain checks in the container it just built}---does the
test it set up actually fail, with an import error?---\textbf{with no agent
involved}. A botched candidate costs a retry, never an agent run. Only a
validated $(x^*, E_0^*)$ is handed to the runner. Guards that could only be
checked by running the agent (predicates over the agent's own mid-run outputs)
are \emph{deferred and counted}, not targeted in v1. \emph{(``LLM proposes,
checker disposes'': the generator may be dumb; the validator guarantees the input
is valid before we pay for a run.)}

\section{Component 6: Property checking (correctness of one run)}
\label{sec:properties}

\paragraph{What ``a bug'' means.} A run is buggy if it \textbf{violates a
property we fixed in advance}---a mechanical automaton rejecting the run, or a
concrete command failing in the final container---\textbf{not} if a model
``thought it went badly.'' Properties are checked over the run's episode
sequence and its final environment.

\paragraph{Two axes: what a property reads, and how it is written.}
A property is classified along two independent axes.

\emph{What it reads:}
\begin{itemize}[leftmargin=1.5em, itemsep=1pt]
  \item \textbf{Trace-structural} --- checkable \emph{directly on the episode
  sequence} (equivalently, on the path through the behavior tree), with no extra
  execution: ordering and presence patterns like ``a \texttt{lint} episode occurs
  before the final episode,'' ``a \texttt{test} episode with a passing result
  precedes any \texttt{commit} episode,'' ``no episode repeats more than $N$
  times.'' These are temporal patterns over the episodes the tree already holds,
  so they are essentially free to evaluate---walk the path, check the pattern.
  Much of the self-consistency and process-hygiene catalog is of this kind.
  \item \textbf{State-based} --- needs a concrete check run against the
  \emph{final container} (or specific files): ``the build passes,'' ``the target
  test still exists with unchanged assertions,'' ``the requested artifact is
  present.'' These execute a command at checking time.
\end{itemize}

\emph{How it is written:}
\begin{itemize}[leftmargin=1.5em, itemsep=2pt]
  \item \textbf{Fixed-formula properties} --- written once, task-independent,
  evaluated by plain code over the trace and run manifest (no force-push, no
  destructive operation outside the workspace, no pathological repetition,
  terminated within budget). Zero model involvement at authoring \emph{and}
  evaluation time; fully trustworthy. These are the universal process
  invariants, run for every skill and every method.
  \item \textbf{Specification-by-example properties} --- written as a
  \emph{natural-language specification} per skill, then \textbf{compiled per
  concrete $(x, E_0)$ into an executable check---before the run exists}. The
  compiling model sees the prompt and the \emph{built initial environment} (its
  tools, its starting file tree), never the run it will judge: the specification
  (``the skill must not make the target test pass by altering the test rather
  than the code'') is reusable; the compiled check is case-specific and must
  \emph{discover} its concrete targets mechanically (find the test files, grep
  the trace) since the agent's actual behavior is unknown at compile time.
  \textbf{The model runs only at compile time}; the produced check is a
  concrete, inspectable bash script that runs mechanically at evaluation
  time---deterministic, auditable, shipped alongside the bugs. Because checks
  attach to the \emph{case}, every run of that case---under any method---is
  judged by byte-identical scripts, which is itself a fairness control. The
  compile step is a calibrated component (its own error rate); we report bug
  yield split by provenance (fixed vs.\ SBE), the SBE ones carrying a wider
  confidence band.
\end{itemize}

\paragraph{One evaluation mechanism.} Every compiled check is a bash script
executed in the run's \emph{final live container}, which exposes the final
workspace, the full trace (\texttt{/check/trace.jsonl}), and the diff of
everything the agent changed---so trace-structural and state-based checks run
through the same auditable mechanism, and a check is free to combine both (grep
the trace \emph{and} run the build).
The two axes are orthogonal: a fixed-formula property can be trace-structural
(``test-before-commit'') or state-based (``build passes''); an SBE property
likewise (``a backup episode precedes modification,'' trace-structural, vs.\
``the target test's assertions are unchanged,'' state-based).

\paragraph{The property catalog ($\sim$18--22 properties, six categories).}
\begin{enumerate}[leftmargin=1.5em, itemsep=1pt]
  \item \textbf{Test integrity} (mostly SBE; the reward-hacking heart): never
  pass the target test by editing the test; never weaken/delete an assertion;
  never add a mock that bypasses the asserted behavior; never override
  equality/comparison to force a pass; never insert \texttt{sys.exit(0)} /
  \texttt{conftest} patches that short-circuit the harness; never hardcode
  expected outputs.
  \item \textbf{Outcome integrity}: never commit/finalize with failing tests
  (\emph{fixed}); leave no dirty repo state unless asked (\emph{fixed});
  the artifact satisfies what the prompt requested (\emph{SBE}).
  \item \textbf{Reversibility \& authorization} (mostly fixed): no force-push
  without confirmation; no destructive/irreversible op without a preceding
  confirmation; verify-before-acting on side-effecting calls; stay within the
  skill's permitted tools/scope (e.g.\ designated working directory).
  \item \textbf{Process hygiene} (fixed): terminate / respect a step budget; no
  pathological repetition; no unnecessary edits when code is already correct;
  never consult the solution/answer-key (e.g.\ reading the fix from git history).
  \item \textbf{Build \& static safety} (fixed, executable): build/lint/compile
  passes in the final state; no newly introduced insecure pattern.
  \item \textbf{Self-consistency} (SBE, from the skill's own docs): whatever the
  \texttt{SKILL.md} advertises holds on the run (``runs the linter before
  finishing'' $\Rightarrow$ a lint episode precedes the final episode).
\end{enumerate}
Not every property applies to every skill; we maintain a per-skill applicability
matrix (a rebasing skill cares about force-push; a fix-the-test skill cares about
test integrity). The fixed core ($\sim$6--8 invariants) is the reusable library
applied to all skills; the rest are selected by relevance.

\paragraph{Reproducibility grading.} A flagged violation is re-run $k{=}3$ times:
3/3 is a genuine skill bug; 1/3 is brittleness (the skill is unreliable here).
Each report carries the violated property, \textbf{the mutated assumption that
produced the input}, a replayable Docker repro, and (for SBE properties) the
generated check that fired.

\paragraph{The honest limitation.} Bug yield is bounded by property strength: a
bug nothing forbids is invisible. This is shared by all assertion-based testing.
We address it empirically by \emph{injecting known violations} (programmatically
producing runs that delete a test, hardcode an output, or force-push) and
reporting the detection rate---turning ``are the properties strong enough?'' into
a measured number rather than a claim. We frame properties as evidence of bugs (a
violation is a real problem), not as a completeness guarantee (``no violation''
$\neq$ ``correct'').

\section{The full loop, assembled}
\label{sec:loop}

\begin{tcolorbox}[colback=blue!4,boxrule=0.5pt]
\small
\textbf{Seed phase.} Run the skill on $\sim$20 seed inputs. Segment, summarize,
and fold each into the tree. Extract guards at every branch. Result: a small
behavior tree and an initial branch frontier.

\smallskip
\textbf{Exploration loop} (until run budget exhausted):
\begin{enumerate}[leftmargin=1.4em, itemsep=1pt, topsep=2pt]
  \item \textbf{Pick} a (branch, mutation) from the frontier by
  \emph{property-guided targeting} (Section~\ref{sec:guards}): among untried,
  $E_0$-decidable mutations, the one most likely to drive a property violation.
  \item \textbf{Mutate} its guard accordingly (negate / target a novel sibling
  value, given all observed siblings).
  \item \textbf{Synthesize} a candidate $(x^*, E_0^*)$ with the generator.
  \item \textbf{Validate} it against the built container (no agent run). Reject
  and retry if invalid.
  \item \textbf{Run} the agent once on the validated input.
  \item \textbf{Fold} the new trace into the tree. Classify:
  \begin{itemize}[leftmargin=1.2em, topsep=1pt]
    \item \emph{Predicted divergence} --- it took the new branch: coverage
    gained, guard corroborated, frontier extended.
    \item \emph{No divergence} --- it reached the branch but behaved as before
    despite the mutated condition: the stated reason was not causal (or the
    mutation did not change what actually mattered). The mutation is marked
    spent and the guard demoted; the run still folds in, so if a genuinely
    different situation was hiding behind the merge, its own divergence
    surfaces as a new branch downstream.
    \item \emph{Path miss} --- it never reached the branch: an earlier guard
    failed; recurse on that.
  \end{itemize}
  \item \textbf{Check properties} on the new run; emit any violations as bug
  reports (with $k{=}3$ regrade).
\end{enumerate}

\smallskip
\textbf{Outputs.} (i)~the behavior tree + coverage report (what of the skill is
explored / unexplored); (ii)~the bug reports (each with its violated property,
the mutated assumption, and a replayable repro); (iii)~per-component agreement
numbers for the model-driven steps.
\end{tcolorbox}

\paragraph{Budget.} $\sim$100--150 agent runs per skill per method. The expensive
model runs once per loop iteration (step~5); everything else is the cheap model
or code. A typical campaign of $\sim$20 skills is a few thousand agent runs---days
on parallel containers.

\section{Baselines: the three-rung ladder}
\label{sec:baselines}

All baselines share \tool's runner, Docker environments, run budget, episode
abstraction (where they need a behavioral signal), and---critically---the
\textbf{same property checker} (so any difference measures \emph{test
generation}, not detection). Each rung adds exactly one ingredient.

\begin{enumerate}[leftmargin=1.5em, itemsep=3pt]
  \item \textbf{Floor --- random mutation.} A model mutates seed prompts and seed
  environments at random; no feedback, every mutant is run. (This is the standard
  black-box fuzzing baseline; it is also structurally the black-box baseline used
  by VeriGrey.)
  \item \textbf{Greybox --- VeriGrey-inspired tool-sequence feedback.} Ports
  VeriGrey's feedback and scheduling \emph{verbatim}: the \emph{novelty of the
  tool-call sequence} decides which mutants join the seed corpus, and energy is
  assigned exactly as published ($+1$ each for a new tool, a new transition, a
  new sequence; energy = offspring count). Mutation is an LLM rewrite of the
  chosen seed's task/environment conditioned on the tool sequence it produced
  (VeriGrey's mutation scheme minus its injection-specific context bridge, which
  has no analogue outside injection testing); its injection oracle is replaced by
  the shared property checker. One adaptation parameter is genuinely free: the
  \emph{tool-event label granularity}. VeriGrey labels events by tool name +
  arguments, which is meaningful over AgentDojo's small argument spaces but
  degenerates on coding agents, where raw arguments are near-unique and would
  mark every run novel. We therefore define three schematization levels (name
  only; name + command head / file extension; name + normalized path), report
  the sensitivity, and use the \textbf{best-performing level} for the headline
  comparison---the baseline gets its empirically best configuration.
  \textbf{This rung is also the ``no-reasoning, no-intent-layer'' ablation}: it
  explores by raw tool-sequence novelty, with no episodes, no reasoning-derived
  guards.
  \item \textbf{\tool.} Episodes + reasoning-derived guard mutation + the behavior
  tree.
\end{enumerate}

\paragraph{What the comparison shows.} Floor~$\to$~Greybox isolates the value of
\emph{any} behavioral feedback over blind mutation. Greybox~$\to$~\tool isolates
the value of \emph{the episode abstraction and reasoning-guard mutation} over raw
tool-sequence novelty---this is the headline claim, and it is measured against a
real published mechanism rather than a strawman. \emph{Optional security slice:}
adopt VeriGrey's own injection oracle on a subset to show \tool is not worse on
the task VeriGrey was built for; kept secondary to stay on the correctness story.

\paragraph{Honest caveat.} The greybox rung is a faithful port of VeriGrey's
\emph{feedback idea}, not VeriGrey itself (we removed its injection-specific
parts), so the comparison speaks to ``tool-sequence feedback vs.\ reasoning-guard
mutation for correctness testing,'' not to VeriGrey's security performance.

\section{The two claims and their experiments}
\label{sec:evalplan}

\paragraph{Claim 1 --- finding skill defects.} On the mined skill suite, run the
three rungs under identical budgets and measure, per skill: distinct violated
properties per agent-run budget, runs-to-first-violation, wall-clock-to-first-%
violation, and unique behavioral branches covered. The shared runner and the
per-case byte-identical checks mean any difference is attributable to test
generation. Here a ``bug'' in a skill is concretely: guidance that is ambiguous
or incomplete in a way that lets a faithful agent violate a correctness
property---each report carries the violated property, the mutated assumption
that produced the input, and a replayable Docker repro.

\paragraph{Claim 2 --- improving generated skills.} Most skills today are
LLM-written, which is circular: a model teaching itself what it may or may not
know. We test whether \tool's feedback breaks the circle. Per skill scenario:
$x$ \emph{hidden} tests (prompt + environment), each with $y$ runnable
pass-criteria---authored \textbf{independently} of every improvement loop
(different model or hand-written; disjoint from anything a tester generates), the
leakage control. Conditions: the LLM-generated skill zero-shot; the skill
revised with random-testing feedback; with greybox feedback; with \tool
feedback (bug reports carrying mutated assumptions + the coverage map of
unexercised guidance). \textbf{The reviser prompt is identical across
conditions---only the feedback payload differs}---so a hidden-test pass-rate
difference is attributable to the feedback quality, i.e.\ to the testing
method. Report pass rates zero-shot vs.\ each revision, and pass-rate per unit
of testing spend.

\end{document}
